---
title: "Isogenic model - multiple pairwise median tests"
# author: "Michal Burdukiewicz"
date: "18-09-2018"
output:
  html_document: default
  pdf_document: default
---

### Raw data 

Raw data from file.

```{r, echo=FALSE,fig.height=150,fig.width=30,message=FALSE,warning=FALSE}
library(readxl)
library(reshape2)
library(multcomp)
library(PMCMR)
library(ggbeeswarm)
library(DT)
library(dplyr)
source("plot_tools.R")

median.test <- function(x, y){
    z <- c(x, y)
    g <- rep(1:2, c(length(x), length(y)))
    m <- median(z)
    if(length(unique(z < m)) == 1) {
      list(p.value = 1)
    } else {
      fisher.test(z < m, g)
    }
}


all_prot <- read.csv("all_prot.csv")

all_prot <- all_prot[, -c(4,11)]
raw_dat <- all_prot[, -2]

colnames(raw_dat)[1L:2] <- c("strain", "protein")
colnames(raw_dat)[3L:8] <- c("R1", "R2", "R3", "R4", "R5", "R6")
#dat <- data.frame(raw_dat[, 1L:2], apply(raw_dat[, -c(1, 2)], 2, function(i) i/max(i)))

dat <- raw_dat
dat[[1]] <- factor(dat[[1]])
dat[[2]] <- factor(dat[[2]])

# dat[["strain"]] <- relevel(dat[["strain"]], "Empty")

amdat <- melt(dat, variable.name = "replicate") %>% 
  arrange(strain, protein, replicate) %>% 
  mutate(rep_gr = ceiling(as.numeric(substr(replicate, 2, 2))/2))

# ggplot(mdat, aes(x = replicate, y = protein, fill = value)) +
#   geom_tile(color = "black") +
#   facet_wrap(~ strain) +
#   my_theme

# amdat <- mutate(mdat, value = value/filter(mdat, strain == "Empty")[["value"]]) %>%
#   filter(strain != "Empty") %>%
#   droplevels()

samdat <- group_by(amdat, strain, protein) %>% 
  summarise(value = median(value)) %>% 
  ungroup

ggplot(amdat, aes(x = protein, y = value)) +
  geom_quasirandom() +
  geom_point(data = samdat, aes(x = protein, y = value), color = "red", size = 2, alpha = 0.5) +
  facet_wrap(~ strain, ncol = 5) + 
  my_theme + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete("") +
  scale_y_continuous("Adjusted value")
```

### Median test

Mood's median test: Fligner, M.A., and Rust, S.W. (1982). A Modification of Mood’s Median Test for The Generalized Behrens--Fisher Problem. Biometrika 69, 221–226.

Reason: does not hold the assumption of the similar variance between two samples.

Multiple testing, Benjamini-Hochberg correction. On the plot below as significant are marked comparisons with p-values smaller than 0.05.

```{r, echo=FALSE,fig.height=150,fig.width=30,message=FALSE,warning=FALSE}
test_res <- lapply(unique(amdat[["strain"]]), function(single_strain) {
  path_dat <- filter(amdat, strain == single_strain)
    # group_by(pathoprotein, strain) %>%
    # summarise(value = median(value)) %>% 
    # ungroup
  
  lapply(combn(levels(path_dat[["protein"]]), 2, simplify = FALSE), function(i) {
    dat1 <- filter(path_dat, protein == i[[1]])[["value"]]
    dat2 <- filter(path_dat, protein == i[[2]])[["value"]]

    pval <- median.test(dat1, dat2)[["p.value"]]
        
    # pval <- try(median.test(dat1, dat2)[["p.value"]], silent = TRUE)
    # if(class(pval) == "try-error")
    #   browstrain()
    
    data.frame(Var1 = i[1], med1 = median(dat1), 
               Var2 = i[2], med2 = median(dat2), pval = pval)
  }) %>%
    do.call(rbind, .) %>% 
    # mutate(apval = p.adjust(pval, method = "BH"), signif = apval < 0.05,
    mutate(apval = pval,
           lbl = paste0("m1 = ", round(med1, 2), "\nm2 = ", round(med2, 2)),
           strain = single_strain)
}) %>% 
  do.call(rbind, .) %>% 
  # select(Var1, Var2, value = apval, strain, lbl, pval) %>% 
  select(Var1, Var2, value = pval, strain, lbl) %>% 
  mutate(significant = value < 0.05)

test_res2 <- rbind(test_res,
                   mutate(test_res, Var1a = Var1, Var1 = Var2, Var2 = Var1a) %>% 
                     select(-Var1a)) %>% 
  mutate(Var2 = factor(Var2, levels = levels(Var2)[c(9, 1L:8)]))

ggplot(test_res2, aes(x = Var1, y = Var2, fill = significant, 
                     label = ifelse(value > 0.05, formatC(value, digits = 2, format = "f"),
                                    format(value, digits = 2)))) +
  geom_tile(color = "black") +
  geom_text(size = 5) +
  facet_wrap(~ strain, ncol = 5) + 
  my_theme + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r, echo=FALSE,fig.height=9,fig.width=8,message=FALSE,warning=FALSE}
datatable(test_res, options = list(dom = "Brtip"), filter = "top") %>% 
  formatRound("value", 4)
```

