---
title: "Isogenic model - multiple pairwise median tests"
author: "Michal Burdukiewicz"
date: "5-01-2017"
output: html_document
---

### Raw data 

This is the raw data from file "06.06.2016 isogenic model 5744 for statistics.xlsx".

```{r, echo=FALSE,fig.height=9,fig.width=8,message=FALSE,warning=FALSE}
library(readxl)
library(reshape2)
library(multcomp)
library(PMCMR)
library(ggbeeswarm)
library(DT)
library(dplyr)
source("plot_tools.R")

median.test <- function(x, y){
    z <- c(x, y)
    g <- rep(1:2, c(length(x), length(y)))
    m <- median(z)
    fisher.test(z < m, g)
}

# raw_dat <- read_excel("06.06.2016 isogenic model 5744 for statistics.xlsx")
all_prot <- read.csv("all_prot.csv")
all_prot <- all_prot[, -c(4,11)]
all_prot <- all_prot[, -2]
raw_dat <- all_prot
# raw_dat <- all_prot[1:920,]
# raw_dat <- all_prot[1:1150,]   # This still works
# raw_dat <- all_prot[1:1380,]   #SU2 is broken?
# raw_dat <- all_prot[-(1151:1610),]  #RNAse and BSA
# colnames(raw_dat)[1L:2] <- c("ser", "type")
colnames(raw_dat)[1L:2] <- c("ser", "type")
colnames(raw_dat)[3L:8] <- c("R1", "R2", "R3", "R4", "R5", "R6")
#dat <- data.frame(raw_dat[, 1L:2], apply(raw_dat[, -c(1, 2)], 2, function(i) i/max(i)))
dat <- raw_dat
dat[[1]] <- factor(dat[[1]])
dat[[2]] <- factor(dat[[2]])

# dat[["ser"]] <- relevel(dat[["ser"]], "Empty")

mdat <- melt(dat, variable.name = "replicate") %>% 
  arrange(ser, type, replicate) %>% 
  mutate(rep_gr = ceiling(as.numeric(substr(replicate, 2, 2))/2))

# ggplot(mdat, aes(x = replicate, y = type, fill = value)) +
#   geom_tile(color = "black") +
#   facet_wrap(~ ser) +
#   my_theme



# amdat <- mutate(mdat, value = value/filter(mdat, ser == "Empty")[["value"]]) %>%
#   filter(ser != "Empty") %>%
#   droplevels()

amdat <- mdat

samdat <- group_by(amdat, ser, type) %>% 
  summarise(value = median(value)) %>% 
  ungroup

ggplot(amdat, aes(x = type, y = value)) +
  geom_quasirandom() +
  geom_point(data = samdat, aes(x = type, y = value), color = "red", size = 2, alpha = 0.5) +
  facet_wrap(~ ser, ncol = 2) + 
  my_theme + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete("") +
  scale_y_continuous("Adjusted value")
```

### Median test

Mood's median test: Fligner, M.A., and Rust, S.W. (1982). A Modification of Mood’s Median Test for The Generalized Behrens--Fisher Problem. Biometrika 69, 221–226.

Reason: does not hold the assumption of the similar variance between two samples.

Multiple testing, Benjamini-Hochberg correction. On the plot below as significant are marked comparisons with p-values smaller than 0.001.

```{r, echo=FALSE,fig.height=13,fig.width=8,message=FALSE,warning=FALSE}
test_res <- lapply(unique(amdat[["ser"]]), function(single_ser) {
  path_dat <- filter(amdat, ser == single_ser) 
    # group_by(pathotype, strain) %>% 
    # summarise(value = median(value)) %>% 
    # ungroup
  
  lapply(combn(levels(path_dat[["type"]]), 2, simplify = FALSE), function(i) {
    dat1 <- filter(path_dat, type == i[[1]])[["value"]]
    dat2 <- filter(path_dat, type == i[[2]])[["value"]]
    pval <- median.test(dat1, dat2)[["p.value"]]
    data.frame(Var1 = i[1], med1 = median(dat1), 
               Var2 = i[2], med2 = median(dat2), pval = pval)
  }) %>% 
    do.call(rbind, .) %>% 
    mutate(apval = p.adjust(pval, method = "BH"), signif = apval < 0.05,
           lbl = paste0("m1 = ", round(med1, 2), "\nm2 = ", round(med2, 2)),
           ser = single_ser)
}) %>% 
  do.call(rbind, .) %>% 
  select(Var1, Var2, value = apval, ser, lbl) %>% 
  mutate(significant = value < 0.001)

test_res2 <- rbind(test_res,
                   mutate(test_res, Var1a = Var1, Var1 = Var2, Var2 = Var1a) %>% 
                     select(-Var1a)) %>% 
  mutate(Var2 = factor(Var2, levels = levels(Var2)[c(9, 1L:8)]))

ggplot(test_res2, aes(x = Var1, y = Var2, fill = significant, 
                     label = ifelse(value > 0.05, formatC(value, digits = 2, format = "f"),
                                    format(value, digits = 2)))) +
  geom_tile(color = "black") +
  geom_text(size = 2.5) +
  facet_wrap(~ ser, ncol = 2) + 
  my_theme + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r, echo=FALSE,fig.height=9,fig.width=8,message=FALSE,warning=FALSE}
datatable(test_res, options = list(dom = "Brtip"), filter = "top") %>% 
  formatRound("value", 4)
```

